---
title: "Example using Berry, Levinson, Pakes (1995)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example using Berry, Levinson, Pakes (1995)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = requireNamespace("R.matlab")
)
```

# Introduction
We revisits the empirical example of Chernozhukov, Hansen, and Spindler (2015)
(CHS2015, hereafter), which extends the instruments of Berry, Levinson,
and Pakes (1995) (BLP1995, hereafter) and applies an instrument selection
procedure based on the Lasso. We consider the same instrument extension
and apply a selection of ensemble procedures that combines conventional
linear estimators with computational alternatives including Lasso-based
approaches and random forests.

```{r setup}
library(ddml)
```

# Data construction

Automobile market data from Berry, Levinsohn, Pakes (1995) as retrived from
the reproduction excercise in Chernozhukov, Hansen, and Spindler (2015). 
[link](https://www.aeaweb.org/articles?id=10.1257/aer.p20151022)

```{r}
library(R.matlab)
df <- readMat("data/BLP_data.mat")

# Create a single dataframe containing all variables
df$cdindex <- df$own.dummies <- NULL
BLP95 <- data.frame(matrix(unlist(df), nrow=2217, byrow=F))
colnames(BLP95) <- names(df)
BLP95$model.name <- readMat("data/BLP_data_str.mat")$model.name
BLP95 <- subset(BLP95, select = -c(const, model.id, product))
```

The BLP1995 data is included as an exemplary dataset with the ddml package
under the name `BLP95`.

```{r}
nobs <- length(BLP95$id)
y <- as.matrix(log(BLP95$share) -  log(BLP95$outshr))
D <- as.matrix(BLP95$price)
x1 <- as.matrix(cbind(1, BLP95[, c("hpwt", "air", "mpd", "space", "price")]))
colnames(y) <- "share"
colnames(D) <- "price"
```

We now construct the BLP1995 instruments, which are sums of product
characteristics (excluding price and other potentially endogenous
variables) of other products offered by the firm as well as over
competing firms. The exact instrument specification here follows the
approach of CHS2015.

```{r}
X_ <- x1[, 1:5]; ncol_X <- ncol(X_) # exclude price
sum_other <- sum_rival <- matrix(0, nobs, 5)
for (i in 1:nobs) {
  other_ind <- (BLP95$firmid == BLP95$firmid[i]) &
    (BLP95$cdid == BLP95$cdid[i]) & (BLP95$id != BLP95$id[i])
  rival_ind <- (BLP95$firmid != BLP95$firmid[i]) &
    (BLP95$cdid == BLP95$cdid[i])
  sum_other[i, ] <- colSums(X_[other_ind, , drop = FALSE])
  sum_rival[i, ] <- colSums(X_[rival_ind, , drop = FALSE])
}#FOR
Z_ <- cbind(sum_other, sum_rival); ncol_Z <- ncol(Z_)
```

In addition, we may be interested in extending the list of instruments. The below 
constructs the additional instruments considered in CHS2015.

```{r}
tu = BLP95$trend/19;
mpdu = BLP95$mpd/7;
spaceu = BLP95$space/2;
XL_ <- as.matrix(cbind(1, BLP95[, c("hpwt", "air")], mpdu, spaceu, tu,
                       BLP95$hpwt^2, BLP95$hpwt^3, mpdu^2, mpdu^3,
                       spaceu^2, spaceu^3, tu^2, tu^3, BLP95$hpwt *
                         BLP95$air,  mpdu * BLP95$air, spaceu *
                         BLP95$air, tu * BLP95$air, BLP95$hpwt *
                         mpdu, BLP95$hpwt * spaceu, BLP95$hpwt * tu,
                       mpdu * spaceu,  mpdu * tu, spaceu * tu))
ncol_XL <- ncol(XL_)
sum_otherL <- sum_rivalL <- matrix(0, nobs, 24)
for (i in 1:nobs) {
  other_ind <- (BLP95$firmid == BLP95$firmid[i]) &
    (BLP95$cdid == BLP95$cdid[i]) & (BLP95$id != BLP95$id[i])
  rival_ind <- (BLP95$firmid != BLP95$firmid[i]) &
    (BLP95$cdid == BLP95$cdid[i])
  sum_otherL[i, ] <- colSums(XL_[other_ind, , drop = FALSE])
  sum_rivalL[i, ] <- colSums(XL_[rival_ind, , drop = FALSE])
}#FOR
ZL_ <- cbind(sum_otherL,sum_rivalL); ncol_ZL <- ncol(ZL_)
```

# Estimating the Flexible Partially Linear IV Model with CV-Lasso

```{r}
# Base learner
learner <- list(what = mdl_glmnet)

# DDML IV. We consider cross-residiualization across 10 sample folds here.
lasso_fit <- ddml_fpliv(y, D = D,
                           Z = ZL_, X = XL_,
                           learners = learner,
                           sample_folds = 3,
                           silent = T)
lasso_fit$coef
```
# Estimating the Flexible Partially Linear IV Model with Stacking


#' A more ambitious approach may consider a variety of computational models for
#'     the estimation of the first and second stage. Here, we consider a
#'     combination of linear models, including unpenalized, Lasso and Ridge
#'     regression, random forests, generalized random forests
#'     (Athey et al., 2019), and boosted trees, using either the inital set of
#'     instruments as in BLP1995 or the extended set as in CHS2015. As before,
#'     we specify  seperate first and second stage model sets (although this is
#'     not strictly necessary).

#' Combine the data. It's convinient to specify sets of indices for easier
#'     column selection in the model specification. When using random forests,
#'     it's important that variables have columnames.
```{r}
X_c <- cbind(X_, XL_); colnames(X_c) <- c(1:ncol(X_c))
Z_c <- cbind(Z_, ZL_); colnames(Z_c) <- c(1:ncol(Z_c))
set_X <- 1:ncol(X_); set_XL <- setdiff(c(1:ncol(X_c)), set_X)
set_Z <- 1:ncol(Z_); set_ZL <- setdiff(c(1:ncol(Z_c)), set_Z)
```



# Estimating the Flexible Partially Linear IV Model with Short-Stacking

```{r}
# Base learners
learners <- list(list(fun = ols,
                       assign_X = set_X,
                       assign_Z = set_Z),
                  list(fun = mdl_glmnet,
                       args = list(alpha = 1),
                       assign_X = set_XL,
                       assign_Z = set_ZL),
                  list(fun = mdl_glmnet,
                       args = list(alpha = 0),
                       assign_X = set_XL,
                       assign_Z = set_ZL))

# DDML IV. We consider cross-residiualization across 10 sample folds here.
stacking_fit <- ddml_fpliv(y, D = D,
                           Z = Z_c, X = X_c,
                           learners = learners,
                           ensemble_type = c("ols", "nnls", "nnls1",
                                             "singlebest", "average"),
                           shortstack = T,
                           sample_folds = 3,
                           silent = T)
stacking_fit$coef
```
